### 2023-09-21 project-plan.Rmd
Comment on project plan

---
title: "Performance of prediction models for opportunities for improvment in trauma care among women compared to men and across patient cohorts"
author: "Anna Thoss"
bibliography: bibliography.bib
csl: vancouver.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "out") })
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- The project plan should be 3-4 pages long and written in
Rmarkdown (like this document). Remove all comments in italic when you 
use this document as a template for your own project plan. -->

<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->

<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->

# Introduction

Traumatic injuries<!-- Call it just trauma--> are a significant public
health concern worldwide, accounting for a substantial portion of
morbidity and mortality rates, being the cause of more than 5 million
fatalities each year [@WHO2014]<!-- This is an old reference, most
current figure is 4.3 million. Use:
@article{injuries2020,
  title =	 {Injuries—Level 1 cause},
  volume =	 {396},
  ISSN =	 {0140-6736},
  journal =	 {The Lancet},
  publisher =	 {Elsevier BV},
  author =	 {{GBD 2019 Diseases and Injuries Collaborators}},
  year =	 {2020},
  month =	 {Oct},
}
-->.  Trauma, and care of patients
with traumatic injuries, have long been a significant target point for
healthcare improvement programs [@WHO2009].  Multidisciplinary
mortality and morbidity conferences form the cornerstone of
initiatives dedicated to enhancing the quality of trauma care and,
consequently, improving patient outcomes [@Santana2014].

At mortality and morbidity conferences, representatives from different
professions and specialties adjacent to trauma care come together to
discuss care provided to a specific patient and compare it to care
given under optimal conditions. The findings of such a conference are
whether opportunities for improvement (OFI) can be found in any given
case.  Examples of opportunities for improvement<!-- You can use OFI
here, you abbreviated it in the previous sentence--> may include lack
of resources and management errors [@Lazzara2020].

Audit filters are applied to determine possible events in patient care
that are associated with unfavourable outcomes. The occurrence of such
an event triggers a review process, and when appropriate leads to the
implementation of corrective measures to address systematic errors and
feedback to individual practitioners [@McDermott_1994]. The use of
audit filters has its drawbacks, being associated with high rates of
false positives. [Saknar tydlig källa här]<!-- Consider changing the
order of these paragraphs, so that you introduce audit filters before
mortality and morbidity conferences, considering that audit filters
are used to select patients for those conferences-->

To enhance the precision of the selection process, there have been
efforts to introduce trauma mortality prediction models. Nevertheless,
the performance of these models has been unsatisfactory
[@Ghorbani2018] [@Heim_2016]<!-- When you cite more than one
reference, use [@Ghorbani2018; @Heim_2016]--><!-- Elaborate, how was
the performance unsatisfactory?--> . This is likely attributed to the
fact that<!-- "attributed to the fact that" can be just "because"-->
these models were originally developed with a focus on predicting
mortality rather than morbidity or instances of care failures within
the realm of trauma care. Machine learning-based prediction models
have been found to outperform conventional methods to predict
opportunities for improvement. *Hur gör jag med källhänvisning här?,
referera till preprint artikel?*<!-- You can do that-->

Men make up a disproportionally large group compared to women when it
comes to trauma. Each year, almost twice as many men die due to
traumatic injury when compared to women, with the distribution of
different types of injury varying by demographic and region
[@WHO2014].  It is important, that due to the differences between
these groups that the model for prediction has similar performance in
these subgroups.  Furthermore, it is important to establish if trained
models perform equally well across different cohorts within traumatic
injuries.<!-- Elaborate on these cohorts-->

# Aim

To assess the difference in performance of prediction models for
opportunities for improvement in trauma care among women compared to men
and across clinical trauma cohorts.

# Methods
<!-- I would find this section easier to read if you followed some
reporting guideline, like TRIPOD--> This is a registry-based study,
encompassing all trauma patients enlisted in both the Karolinska
University Hospital trauma registry and the trauma care quality
database spanning from 2012 to 2022. The Karolinska University
Hospital in Solna is equivalent to a level 1 trauma centre and manages
approximately 1,500 acute trauma patients each
year[@Ghorbani2018][Swetrau, rapport]. The trauma registry comprises
all patients admitted to the hospital with trauma team activation,
irrespective of their Injury Severity Score (ISS), as well as patients
with an ISS exceeding 9. Included in the registry is data on vital
signs, injuries, interventions, times, and patient demographics.

The trauma registry encompasses approximately 12,000 patients treated
from 2012 to 2022. The trauma care quality database in turn is a subset
of the trauma registry, including about 6000 patients selected for
review in the years 2013-2022.

The outcome of this study is opportunities for improvement (OFI), which
are identified through meetings of the multidisciplinary review board
during mortality and morbidity conferences. An OFI is a consensus
decision made during a conference and includes proposals for corrective
actions. These in turn are recorded in the trauma care quality database.
The trauma care quality database also includes data relevant to
mortality and morbidity conferences, such as audit filters. The
mortality and morbidity conferences at Karolinska University Hospital
involve professionals from various specialties within trauma care,
including surgery, neurosurgery, orthopedics, anesthesia, intensive
care, nursing, and radiology. Cases considered for a conference go
through multiple rounds of review. Mortality leads to direct inclusion.
Between 2013 and 2017, non-fatal poor outcomes were identified through
reviews conducted by a specialized trauma nurse to identify potential
OFI. Starting in 2017, the process was formalized, involving an initial
review by a specialized trauma nurse along with the application of audit
filters. All cases that were flagged either during the initial nurse
review or by the audit filters underwent a secondary review conducted by
two specialized nurses. If the second review identified potential areas
for improvement (OFI), the case was marked for discussion in a
multidisciplinary conference.

All patients reviewed for possible opportunities for improvement between
the years 2013 and 2021 are included in the study, with the exception of
patients under the age of 15 due to differing clinical pathways in this
subgroup.

The outcome of this study will be defined as "Yes, there is a
significant difference in models prediction of OFI in different cohorts"
or "No, there is no significant difference in models prediction of OFI
in different cohorts." Data for this will be extracted from the trauma
care quality database and run through different previously trained
machine learning models. *osäker på om outcome här är hållbart
formulerad*<!--Good enough for now-->

All variables from the trauma registry will be considered as potential
predictors in our analysis<!-- Note that the models are already
developed. So you should write about what have already been done,
i.e. these variables are already incorporated into the
models-->. These variables include information spanning the
pre-hospital, initial care, and subsequent in-hospital phases, which
include initial vital signs, timing, and types of procedures and
interventions, duration and level of care, injury details, mechanisms
of injury, injury types, and standard demographic information. This
array of predictors comprises both continuous and categorical
variables, ultimately resulting in the final models utilizing a total
of 45 predictors. (infoga bilaga) All available data will be used,
with a sample size of 6310 patients. This is due to previously
developed prediction models being based on all available data, as well
as including all 45 predictors regardless of the learner used for the
development of the model.

The models have been developed and trained on the data set previously
to this study<!-- Add preprint citation-->. Eight models are included,
all being developed using the Tidymodels framework (). These include
logistic regression (LR), random forest (RF), decision tree (DT),
support vector machine with a radial basis kernel (SVM), XGBoost,
LightGBM, CatBoost, and k-nearest neighbor (k-NN).

The different models' performance were assessed in the following cohorts:

1.  Blunt multi-system trauma: Blunt trauma with AIS ≥ 3 in at least two
    of the following AIS body regions: head, face, neck, thorax,
    abdomen, spine, or upper and lower extremities.

2.  Isolated severe TBI: Injury isolated to the area of the brain with
    AIS \> 2 and:

    a.  pre-hospital GCS of \< 9

        or

    b.  Pre- or in-hospital intubation.

3.  Blunt multi-system trauma without TBI: Blunt trauma with AIS \> 2
    and injuries in at least two of the following AIS body regions:
    head, face, neck, thorax, abdomen, spine, or upper and lower
    extremities and no injury to the brain with AIS \>2 and:

    a.  pre-hospital GCS of \< 9

        or

    b.  Pre- or in-hospital intubation.

4.  Blunt multi-system trauma with TBI: Blunt trauma with AIS \> 2 and
    injuries in at least two of the following AIS body regions: head,
    face, neck, thorax, abdomen, spine, or upper and lower extremities
    and with injury to the brain with AIS \>2 and:

    a.  pre-hospital GCS of \< 9

        or

    b.  Pre- or in-hospital intubation

Along with these cohorts, all machine learning models will be assessed
in regard to differences between male and female sex in their
prediction of OFI.

We will employ a 95% confidence level and a 5% significance level for
our analysis, with all statistical analysis conducted using R.

# Project Update and Timeline

![**The Gantt chart will be displayed here when you place it in the
assets folder under "project-gantt.png"**](assets/project-gantt.png)

Gantt chart over the project timeline.

# Back-up Plan

Currently, there is no need to develop a thorough backup plan since all
the necessary data for the statistical analysis has been collected and
is accessible.

# References
=======

Introduction
============

*This is part A) in the project plan [sv: projektredogörelse]. *

*The background/introduction should summarise the scope of the problem,
what is known about the problem, what is not known, what specific
knowledge gap the study is supposed to fill and why this is
important. It should end be stating the aim of the study. The
background should be 4-5 paragraphs long and each paragraph should be
between 3-5 sentences. Strive to make the paragraphs about the same
length.*

*bibliography.bib includes an example reference. Add additional
references in that file as bibtex entries and then cite in text using:
"[@exampleKey9999]". One can convert DOIs' to bibtex entries using 
https://www.doi2bib.org/. If you use Zotero you can simply export 
entries to bibtex `Ctrl-Shift-C`, however you need to first setup 
Quick Copy to export to bibtex (https://www.zotero.org/support/creating_bibliographies#quick_copy)*

Methods
=======

*Refer to the appropriate reporting guideline for details. If you are
developing, updating or validating a clinical prediction model then
use
[TRIPOD](https://www.equator-network.org/reporting-guidelines/tripod-statement/). If
you are conducting an observational study, for example a cohort or
case control study in which you assess associations between some
exposure and an outcome then use
[STROBE](https://www.equator-network.org/reporting-guidelines/strobe/).*

Project Update and Timeline
===========================

*This is part B) in the project plan [sv: projektredogörelse]. *

*Save the image generated from the Gantt chard spreadsheet into the assets folder, rename it to project-gantt.png*

![**The Gantt chart will be displayed here when you place it in the assets folder under "project-gantt.png"**](assets/project-gantt.png)

Back-up Plan
============

*This is part C) in the project plan [sv: projektredogörelse]*

References
==========

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->